---
title: "Assignment 2 - Getting Started with R (4 pts)"
author: "John Hope (jah9kqn)"
date: "Due Date: 11:59pm, Jan 26"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
---

<style>
h1.title {
  font-size: 30px;
}
h1 {
  font-size: 26px;
}
h2 {
  font-size: 22px;
}
h3 { 
  font-size: 18px;
}

</style>

# Read the csv file "inc_real.sav" (contains real income data) into R and then check the data by doing the following:

- Print the first 6 or 10 rows of the data frame
- Use `dim()` to determine how many rows and columns the data frame has
- Get the variable names (use `names()`),
- Determine the type of variables (numerical, factor, …), i.e., the structure of the data frame (`str()`).

```{r, warning=FALSE, message=FALSE}
library(foreign)
inc_real <- read.spss('/Users/johnhope/Desktop/DS3003/Data/inc_real.sav')

inc_real <- as.data.frame(inc_real) #converting to a data frame

head(inc_real) #printing first 6 rows
```
We see the first 6 rows of the data

```{r}
dim(inc_real)
```
The data has 1271 rows and 8 columns

```{r}
names(inc_real)

str(inc_real)
```
Above we see the variable names and their associated types

    
# Get the summary statistics for the variables in the data frame.

```{r}
summary(inc_real)
```

# Generate the following sequences using seq() and rep(). 

- Each sequence should have a length of 20 (i.e., 20 numbers), only the first 12 numbers are shown below.

   - 1 0 1 0 1 0 1 0 1 0 1 0 ….
   - 1 1 0 0 1 1 0 0 1 1 0 0 ….
   - 0 3 6 9 0 3 6 9 0 3 6 9 ….

```{r}
rep(c(1,0),10)
rep(rep(1:0, each = 2), 5)
rep(seq(0, 9, by=3), 5)
```

# Plot a histrogram with the income variable from data. 

```{r}
hist(inc_real$income)
```
